Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-14 WholePlotTrajectories.Rmd
conclusions

In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss. We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy. Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in its aspect of taxonomic and functional diversity and time after disturbance, and (iii) we analyzed community resilience and time to recovery.

@@

# Conclusions

Our study highlighted the combination of deterministic and stochastic processes determining the succession of recruited species and shaping community response to disturbance. 
The deterministic recruitment of pioneers shaped functional trajectories convergent in the functional space, while dispersal limitations shaped stochastic taxonomic trajectories maintaining community divergence in taxonomic composition.
In accordance with the IDH, community taxonomic diversity was maximised for intermediate disturbance intensity, before reaching an intensity threshold beyond which the taxonomic richness decreased and the taxonomic evenness remained persistently higher.
Conversely, functional trajectories proved decoupled from the taxonomic ones and did not follow the hypothesis of the IDH as no distinct trajectories emerged with increasing disturbance intensity. 
The decoupling between taxonomic and functional disturbance was explained by a decrease of community functional redundancy mitigating the functoinal impact of disturbance.
The resilience of tropical forests, defined in terms of recovery to pre-disturbance state, proved tangible but requiring several decades. 
Still, the disturbance impact on community redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

>>>>

In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss. We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy. Specifically, (i) we defined taxonomic and functional post-disturbance trajectories and examined the underlying ecological process, (ii) we discussed the scope of the IDH regarding the different facets of community diversity and regarding time after disturbance, and (iii) we analyzed community resilience and time to recovery.

@@

# Conclusion

Community compositon and diversity post-disturbance trajectories were shaped by the recruitment of pioneers, identical for all communities and disturbance intensity.
However, taxonomic and functional composition trajectories were decoupled: while functional trajectories were similar in the functional space and converged towards the recovery of the same initial state, taxonomic trajectories displayed initial composition differences maintained along time.
This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
The disturbance impact was contrasted as well between functoinal and taxonomic diversity.
While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
The Intermediate Disturbance Hypothesis was then validated regarding community taxonomic response but inferred regarding the functional one.
Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance when the recruitment of early- and late-successional species were balanced, that was around 25 years after disturbance.
Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied on the lottery, random recruitment of rare species. 
Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

### 2019-02-14 WholePlotTrajectories.tex
conclusions

In this study, we monitored over 30 years the response of 75 ha of
Neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
large functional traits database encompassing major leaf, stem and
life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy.
Specifically, (i) from post-disturbance trajectories we examined
community taxonomic and functional response, (ii) then we discussed the
scope of the IDH in its aspect of taxonomic and functional diversity and
time after disturbance, and (iii) we analyzed community resilience and
time to recovery.

@@

\section{Conclusions}\label{conclusions}

Our study highlighted the combination of deterministic and stochastic
processes determining the succession of recruited species and shaping
community response to disturbance. The deterministic recruitment of
pioneers shaped functional trajectories convergent in the functional
space, while dispersal limitations shaped stochastic taxonomic
trajectories maintaining community divergence in taxonomic composition.
In accordance with the IDH, community taxonomic diversity was maximised
for intermediate disturbance intensity, before reaching an intensity
threshold beyond which the taxonomic richness decreased and the
taxonomic evenness remained persistently higher. Conversely, functional
trajectories proved decoupled from the taxonomic ones and did not follow
the hypothesis of the IDH as no distinct trajectories emerged with
increasing disturbance intensity. The decoupling between taxonomic and
functional disturbance was explained by a decrease of community
functional redundancy mitigating the functoinal impact of disturbance.
The resilience of tropical forests, defined in terms of recovery to
pre-disturbance state, proved tangible but requiring several decades.
Still, the disturbance impact on community redundancy cautioned against
the risks of infrequent species loss and the persistence of
disturbance-specific communities \citep{Herault2018}.

>>>>

In this study, we monitored over 30 years the response of 75 ha of
Neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
large functional traits database encompassing major leaf, stem and
life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy.
Specifically, (i) we defined taxonomic and functional post-disturbance
trajectories and examined the underlying ecological process, (ii) we
discussed the scope of the IDH regarding the different facets of
community diversity and regarding time after disturbance, and (iii) we
analyzed community resilience and time to recovery.

@@

\section{Conclusion}\label{conclusion}

Community compositon and diversity post-disturbance trajectories were
shaped by the recruitment of pioneers, identical for all communities and
disturbance intensity. However, taxonomic and functional composition
trajectories were decoupled: while functional trajectories were similar
in the functional space and converged towards the recovery of the same
initial state, taxonomic trajectories displayed initial composition
differences maintained along time. This decoupling was explained by a
decrease of the functional redundancy which mitigated the functional
impact of disturbance. The disturbance impact was contrasted as well
between functoinal and taxonomic diversity. While the functional
trajectories remained similar whenever the disturbance intensity,
taxonomic trajectories were markedly different above a threshold of
20-25\% AGB removed for which taxonomic richness was maximized and
taxonomic eveness remained resilient. The Intermediate Disturbance
Hypothesis was then validated regarding community taxonomic response but
inferred regarding the functional one. Taxonomic diversity was maximized
following a disturbance of intermediate intensity, that was the
threshold of 25\% AGB removed, and at an intermediate time after
disturbance when the recruitment of early- and late-successional species
were balanced, that was around 25 years after disturbance. Whenever the
disturbance intensity, community resilience (in terms of recovery of the
pre-disturbance state) was tangible but required several decades and
relied on the lottery, random recruitment of rare species. Although
resilient, the functional redundancy was lower for more than 30 years
after disturbance which cautioned against the risks of infrequent
species loss and the persistence of disturbance-specific communities
\citep{Herault2018}.
